An Embedding Framework for Consistent Polyhedral Surrogates

### 2019-05-23 neurips-19/poly-embed.tex
Addressed Raf's comments on top k

\raf{Jessie: stuff to update / change:}
  \begin{itemize}
  \item Need to mention the link function typically used; don't see it anywhere
  \item You keep saying $L(r,y)$ not $L(r)_y$ for all losses \jessie{fixed}
  \item Let's use our property notation; $\prop{L_2}$ instead of argmins \jessie{fixed}
  \item The story we discussed: YK show $L_2$ is inconsistent.  Our framework lets us ask: with what discrete loss \emph{is} $L_2$ consistent?  (It also assures us that $L_2$ is consistent for something...)
  \item The next part of the story: oh look, $L_2$ embeds something which is $\ell_{top-k}$ plus some other term, so (a) we can see very clearly what $L_2$ ``is doing'', and (b) we can see where to look for distributions yielding inconsistency, namely ones for which we actually prefer to report a set with $|r|<k$.
  \item Domain of $\ell_2$ should not be $\{0,1\}^n$ anymore \jessie{fixed}
  \end{itemize}

@@

  \citet{lapin2016loss, lapin2018analysis}, provide two surrogate losses for measuring top-$k$ error, one of which is shown to be convex, but their consistency is left as an open question.
  \citet{yang2018consistency} show the inconsistency of their losses, and create a new surrogate loss that is consistent for the top-$k$ problem, but it should be noted that Yang and Koyejo's surrogate is not convex.
  They also show that Lapin's surrogates are not consistent, but without too much intuition why.
  Following the notation of Yang, we call Lapin's~\cite{lapin2018analysis} convex surrogate $L_2$, and focus on this.
  Here, we assume $r \subseteq [n]$ and $k < n$ is a constant.

 In order to do this, we want to show that for all $p \in \simplex$, $\inf_{u' \in \reals^n} \inprod{p}{L_2(u')} = \min_{u' \in \{0,1\}^n} \inprod{p}{L_2(u')}$. (Lemma~\ref{lem:top-k-optimal-corners}.) \jessiet{Appendix ref will be broken}
Then, given the set $\R = \{r \subseteq 2^{[n]} : |r| \leq k\}$, we show there is a discrete loss $\ell_2 : \R \to \reals^\Y_+$ that is embedded by $L_2$.  (Lemma~\ref{lem:top-k-surrogate-embeds}.)\jessiet{Appendix ref will be broken}

  We conclude with a concrete example: $k=1$.
  One can see that the $\ell_{top-1} = \ell_{0-1}$, and we characterize the distributions $p \in \simplex$ where $\risk{L_2}(p) = \risk{\ell_2}(p) \neq \risk{\ell_{top-k}}(p)$.

@@

We are given the surrogate 
\begin{align}\label{eq:L-2-surrogate}
L_2(u)_y := \left( \frac{1}{k} \sum_{i=1}^k (u + \ones - e_y)_{[i]} - u_y \right)_+
\end{align}
where $x_{[i]}$ is the $i^{th}$ largest component of $x$.
By Lemma~\ref{lem:top-k-optimal-corners}\jessiet{in the Appendix... remove ref for camera ready version}, we can see that for all $p \in \simplex$, there is some $u \in \{0,1\}^n$ so that $u \in \prop{L_2(u)}$.
With this, we show that $L_2$ actually embeds the discrete loss $\ell_2: \R \to \reals^\Y_+$ given below:
\begin{align}\label{eq:ell-2}
\ell_2(r)_y &= \Ind{y \not\in r} + \frac{|r \setminus \{y\}|}{k}
\end{align}
We show that $L_2$ embeds $\ell_2$ in Lemma~\ref{lem:top-k-surrogate-embeds} \jessiet{in the Appendix... remove ref for camera ready version}.
Moreover, by Proposition~\ref{prop:embed-bayes-risks}, we know $\risk{L_2} = \risk{\ell_2}$.

While $L_2$ given by~\citet{lapin2018analysis} is not consistent with respect to top-$k$ loss, we can at least show a discrete loss $\ell_2$ where $L_2$ is consistent with respect to $\ell_2$, applying Theorem~\ref{thm:embed-implies-calibrated-link}.\jessiet{TODO: Make ref}

Now to see a set of distributions where $\risk{L_2}(p) \neq \risk{\ell_{top-1}}(p)$, we claim that $\vec{0}~\in~\prop{\ell_2} \implies \risk{\ell_2}(p) = 1$ for any $p$ with $\max_y p_y \leq 1/2$.
Additionally $\risk{\ell_{top-1}}(p) = 1 - \max_y p_y < 1$.

First, $\ell(\vec{0}, \bar{p}) = 1$ for all $p \in \simplex$.
Taking the set $\{y\}$ for any $y \in \Y$, the expected loss $\inprod{p}{\ell_2(\{y\})} = 2 (1-p_y) \geq 1$ by assumption on $p$. 
For any $r$ such that $|r| \geq 2$ \jessiet{In general, $|r| \geq k+1$}, we know that  $\ell_2(r)_y \geq 1$ for all $y \in \Y$, so the expected loss much be greater than or equal to $1$.
Therefore, we have that $\vec{0} \in \prop{\ell_2}$ for any $p$ such that $\max_y p_y \leq 1/2$.

>>>>

\raf{Jessie: stuff to update / change:}
  \begin{itemize}
  \item Need to mention the link function typically used; don't see it anywhere \jessie{Fixed, but could be checked.  right under eq 2}
  \item You keep saying $L(r,y)$ not $L(r)_y$ for all losses \jessie{fixed}
  \item Let's use our property notation; $\prop{L_2}$ instead of argmins \jessie{fixed}
  \item The story we discussed: YK show $L_2$ is inconsistent.  Our framework lets us ask: with what discrete loss \emph{is} $L_2$ consistent?  (It also assures us that $L_2$ is consistent for something...) \jessie{Fixed}
  \item The next part of the story: oh look, $L_2$ embeds something which is $\ell_{top-k}$ plus some other term, so (a) we can see very clearly what $L_2$ ``is doing'', and (b) we can see where to look for distributions yielding inconsistency, namely ones for which we actually prefer to report a set with $|r|<k$. \jessie{Fixed with question}
  \item Domain of $\ell_2$ should not be $\{0,1\}^n$ anymore \jessie{fixed}
  \end{itemize}

@@

  \citet{lapin2015top, lapin2016loss, lapin2018analysis}, provide two surrogate losses for measuring top-$k$ error, one of which is shown to be convex, but their consistency is left as an open question.
  \citet{yang2018consistency} show the inconsistency of their losses, and create a new surrogate loss that is consistent for the top-$k$ problem, but it should be noted that Yang and Koyejo's surrogate is not convex.
  They also show that Lapin's surrogates are not consistent, but without too much intuition why.
  Following the notation of Yang, we call Lapin's~\cite{lapin2016loss} convex surrogate $L_2$, and focus on this.
  Here, we assume $k < n$ is a constant.

% In order to do this, we want to show that for all $p \in \simplex$, $\inf_{u' \in \reals^n} \inprod{p}{L_2(u')} = \min_{u' \in \{0,1\}^n} \inprod{p}{L_2(u')}$. (Lemma~\ref{lem:top-k-optimal-corners}.) \jessiet{Appendix ref will be broken... this sentence doesn't really contribute to the story}

We know from Theorem~\ref{thm:poly-embeds-discrete} that every polyhedral loss embeds some discrete loss, and our framework shows us one such loss.
In Equation~\ref{eq:ell-2}, we show a loss embedded by $L_2$ which is similar, but generally not equal to $\ell_{top-k}$.  
So while $L_2$ is not consistent with respect to $\ell_{top-k}$, we can show it is consistent with respect to $\ell_{2}$ by embedding and Theorem~\ref{thm:eps-thick-calibrated}.

%  We conclude with a concrete example: $k=1$.
%  One can see that the $\ell_{top-1} = \ell_{0-1}$, and we characterize the distributions $p \in \simplex$ where $\risk{L_2}(p) = \risk{\ell_2}(p) \neq \risk{\ell_{top-k}}(p)$.

@@

We are given the surrogate 
\begin{align}\label{eq:L-2-surrogate}
L_2(u)_y := \left( \frac{1}{k} \sum_{i=1}^k (u + \ones - e_y)_{[i]} - u_y \right)_+
\end{align}
where $x_{[i]}$ is the $i^{th}$ largest component of $x$.
\jessie{Link added here}
While not explicitly stated, they link the surrogate by the function $\psi(u)$, mapping $u$ to the set of indices for the largest $k$ elements of $u$, with ties broken arbitrarily, but consistently.

By Lemma~\ref{lem:top-k-optimal-corners}\jessiet{in the Appendix... remove ref for camera ready version}, we can see that for all $p \in \simplex$, there is some $u \in \{0,1\}^n$ so that $u \in \prop{L_2}$.
With this, we show that $L_2$ actually embeds the discrete loss $\ell_2: \R \to \reals^\Y_+$ given below:
\begin{align}\label{eq:ell-2}
\ell_2(r)_y &= \Ind{y \not\in r} + \frac{|r \setminus \{y\}|}{k}
\end{align}
We show that $L_2$ embeds $\ell_2$ in Lemma~\ref{lem:top-k-surrogate-embeds} \jessiet{in the Appendix... remove ref}.
%Moreover, by Proposition~\ref{prop:embed-bayes-risks}, we know $\risk{L_2} = \risk{\ell_2}$.

%While $L_2$ given by~\citet{lapin2018analysis} is not consistent with respect to top-$k$ loss, we can at least show a discrete loss $\ell_2$ where $L_2$ is consistent with respect to $\ell_2$, applying Theorem~\ref{thm:embed-implies-calibrated-link}.\jessiet{TODO: Make ref}

Let us re-write $\ell_2(r)_y = \ell_{top-k}(r)_y + \frac{|r \setminus \{y\}|}{k}$.
We can now see pretty clearly what $L_2$ is doing, as well as where to look for distributions yielding inconsistency: those where it is preferable to report a set with $|r| < k$. \jessie{Less than k or greater than 1?}

We now focus on the case where $k=1$; the top-$k$ loss reduces to 0-1 loss in this case.
To show a set of distributions where $\prop{\ell_2}(p) \neq \prop{\ell_{top-k}}(p)$, we claim that $\vec{0}~\in~\prop{\ell_2} \implies \risk{\ell_2}(p) = 1$ for any $p$ with $\max_y p_y \leq 1/2$.
Additionally $\risk{\ell_{top-1}}(p) = 1 - \max_y p_y < 1 \implies \vec{0} \not\in \prop{\ell_{top-k}}(p)$.

First, $\ell(\vec{0}, \bar{p}) = 1$ for all $p \in \simplex$.
Taking the set $\{y\}$ for any $y \in \Y$, the expected loss $\inprod{p}{\ell_2(\{y\})} = 2 (1-p_y) \geq 1$ by assumption on $p$. 
%For any $r$ such that $|r| \geq 2$ \jessiet{In general, $|r| \geq k+1$}, we know that  $\ell_2(r)_y \geq 1$ for all $y \in \Y$, so the expected loss much be greater than or equal to $1$.
Therefore, we have that $\vec{0} \in \prop{\ell_2}(p)$ but $\vec{0} \not\in \prop{\ell_{top-k}}(p)$ for any $p$ such that $\max_y p_y \leq 1/2$.
